---
title: "Cleaning Data"
subtitle: "EMSE 4572: Exploratory Data Analysis"
date: September 14, 2022
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
install.packages('lubridate') # for dealing with dates
install.packages('janitor') # for cleaning data

library(knitr)
library(tidyverse)
library(cowplot)
library(readxl)
library(lubridate)
library(janitor)
library(here)

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.retina = 3)
dplyr.width = Inf

# Load all the data sets:
wildlife_impacts <- read_csv(here::here('data', 'wildlife_impacts.csv'))
milk_production <- read_csv(here::here('data', 'milk_production.csv'))

# Abbreviations data, from: https://www.50states.com/abbreviations.htm
state_abbs <- tibble::tribble(
                  ~state_name,              ~state_abb,
                     "Alabama",             "AL",
                      "Alaska",             "AK",
                     "Arizona",             "AZ",
                    "Arkansas",             "AR",
                  "California",             "CA",
                    "Colorado",             "CO",
                 "Connecticut",             "CT",
                    "Delaware",             "DE",
                     "Florida",             "FL",
                     "Georgia",             "GA",
                      "Hawaii",             "HI",
                       "Idaho",             "ID",
                    "Illinois",             "IL",
                     "Indiana",             "IN",
                        "Iowa",             "IA",
                      "Kansas",             "KS",
                    "Kentucky",             "KY",
                   "Louisiana",             "LA",
                       "Maine",             "ME",
                    "Maryland",             "MD",
               "Massachusetts",             "MA",
                    "Michigan",             "MI",
                   "Minnesota",             "MN",
                 "Mississippi",             "MS",
                    "Missouri",             "MO",
                     "Montana",             "MT",
                    "Nebraska",             "NE",
                      "Nevada",             "NV",
               "New Hampshire",             "NH",
                  "New Jersey",             "NJ",
                  "New Mexico",             "NM",
                    "New York",             "NY",
              "North Carolina",             "NC",
                "North Dakota",             "ND",
                        "Ohio",             "OH",
                    "Oklahoma",             "OK",
                      "Oregon",             "OR",
                "Pennsylvania",             "PA",
                "Rhode Island",             "RI",
              "South Carolina",             "SC",
                "South Dakota",             "SD",
                   "Tennessee",             "TN",
                       "Texas",             "TX",
                        "Utah",             "UT",
                     "Vermont",             "VT",
                    "Virginia",             "VA",
                  "Washington",             "WA",
               "West Virginia",             "WV",
                   "Wisconsin",             "WI",
                     "Wyoming",             "WY",
        "District of Columbia",             "DC",
            "Marshall Islands",             "MH",
         "Armed Forces Africa",             "AE",
       "Armed Forces Americas",             "AA",
         "Armed Forces Canada",             "AE",
         "Armed Forces Europe",             "AE",
    "Armed Forces Middle East",             "AE",
        "Armed Forces Pacific",             "AP"
)

View(wildlife_impacts)
View(milk_production)
View(state_abbs)
```

# 1. Joins

1) Create a data frame called `state_data` by joining the data frames `states_abbs` and `milk_production` and then selecting the variables `region`, `state_name`, `state_abb`. **Hint**: Use the `distinct()` function to drop repeated rows.

```{r}
# Write code here

new_milk <- milk_production %>% 
    select(c('region', 'state')) %>% 
    rename(state_name=state) %>% 
    distinct()
View(new_milk)

state_data <- state_abbs %>% 
    left_join(new_milk, by="state_name")
View(state_data)
```

2) Create the data frame `wildlife_impacts2` by joining the `state_data` data frame to the `wildlife_impacts` data frame, adding the variables `region` and `state_name`.

```{r}
wildlife_impacts2 <- wildlife_impacts %>% 
    rename(state_abb=state)  %>%
    left_join(state_data, by="state_abb")
View(wildlife_impacts2)


wildlife_impacts3 <- wildlife_impacts %>% 
    left_join(state_data, by=c("state_abb"="state"))
View(wildlife_impacts3)

```

# 2. Re-naming and re-typing variables

Read in the `hot_dog_winners.xlsx` file and adjust the variable names and types to match those shown on the screen.

```{r}

# IMPORT 

csv_filepath = here::here("data", "hot_dog_winners.xlsx")
hot_dog_winners <- read_excel(csv_filepath)
View(hot_dog_winners)
print(colnames(hot_dog_winners))

# CLEAN 
# ... consider janitor::clean_names(hot_dog_winners)
# https://www.datasciencemadesimple.com/union-union_all-function-r-using-dplyr-union-data-frames/
# https://stackoverflow.com/questions/34429001/adding-a-column-to-data-frame-and-fill-that-column-with-a-particular-string-usin
# https://stackoverflow.com/questions/9126840/delete-rows-with-blank-values-in-one-particular-column
# https://www.sharpsightlabs.com/blog/case-when-r/

hot_dogs_clean <- hot_dog_winners %>% 
    rename(c(
        'year'='Year',
        
        'male_contestant'='Mens',
        'male_dogs_eaten'='Dogs eaten...3',
        'male_country'='Country...4',
        
        'female_contestant'='Womens',
        'female_dogs_eaten'='Dogs eaten...6',
        'female_country'='Country...7',
    ))

View(hot_dogs_clean)

hot_dog_males <- hot_dogs_clean %>% 
    select(c('year', 'male_contestant', 'male_dogs_eaten', 'male_country')) %>% 
    rename(c(
       'contestant'='male_contestant',
       'dogs_eaten'='male_dogs_eaten',
       'country'='male_country'
    ))
hot_dog_males$gender <- 'MALE'
View(hot_dog_males)

hot_dog_females <- hot_dogs_clean %>% 
    select(c('year', 'female_contestant', 'female_dogs_eaten', 'female_country')) %>% 
    rename(c(
       'contestant'='female_contestant',
       'dogs_eaten'='female_dogs_eaten',
       'country'='female_country'
    ))
hot_dog_females$gender <- 'FEMALE'
View(hot_dog_females)

hot_dog_tall <- union(hot_dog_males,hot_dog_females)
View(hot_dog_tall)

# drop columns that have null contestant
hot_dog_final <- hot_dog_tall[!is.na(hot_dog_tall$contestant), ]

#hot_dog_final$dogs_eaten_clean <- apply(hot_dog_final, 3, parse_number)
#hot_dog_final$new_record <- apply(hot_dog_final, 'dogs_eaten', myFunction)

# split asterisk into a clean col of dogs and a separate col of record
hot_dog_final <- hot_dog_final %>% 
    mutate(
        dogs_eaten_clean = parse_number(dogs_eaten),
        #new_record = case_when(
        #    str_detect(hot_dog_final$dogs_eaten, '\\*') ~ TRUE,
        #    TRUE ~ FALSE
        #)
        new_record = if_else(
            str_detect(hot_dog_final$dogs_eaten, '\\*'), TRUE, FALSE
        )
    ) %>% 
    select(-'dogs_eaten') %>% 
    rename('dogs_eaten'='dogs_eaten_clean')

#case_when(test_score_vector >= 90 ~ 'A'
#          ,test_score_vector >= 80 ~ 'B'
#          ,test_score_vector >= 70 ~ 'C'
#          ,test_score_vector >= 60 ~ 'D'
#          ,TRUE ~ 'F'
#          )

#str_detect('67.5*', '7') #> TRUE

#str_detect('67.5*', '*') #> 
#> Error in stri_detect_regex(string, pattern, negate = negate, opts_regex = opts(pattern)) : 
#> Syntax error in regex pattern. (U_REGEX_RULE_SYNTAX, context=`*`)

# str_detect('67.5*', '\\*') #> TRUE

View(hot_dog_final)
```

# 3. Recoding variables

1) Use `case_when()` to modify the `phase_of_flt` variable in the `wildlife_impacts` data:

- The values `'approach'`, `'arrival'`, `'descent'`, and `'landing roll'` should be merged into a single value called `'arrival'`.
- The values `'climb'`, `'departure'`,  and `'take-off run'` should be merged into a single value called `'departure'`.
- All other values should be called `'other'`.

```{r}
csv_filepath = here::here("data", "wildlife_impacts.csv")
impacts_df <- read_csv(csv_filepath)
View(impacts_df)

arrival_names <- c('approach', 'arrival', 'descent', 'landing roll')
departure_names <- c('climb', 'departure', 'take-off run')

impacts_2 <- impacts_df %>% 
    mutate(
        phase_new = case_when(
            tolower(phase_of_flt) %in% arrival_names ~ 'ARRIVAL',
            tolower(phase_of_flt) %in% departure_names ~ 'DEPARTURE',
            TRUE ~ 'OTHER'
        )
    )
    
View(impacts_2)


```

2) Use the **lubridate** package to create a new variable, `weekday_name`, from the `incident_date` variable in the `wildlife_impacts` data.

```{r}
# Write code here
```

3) Use `weekday_name` and `phase_of_flt` to make this plot of "arrival" and "departure" impacts from **Feb. 2016**.

```{r, echo=FALSE, fig.width=9, fig.height=4.5, fig.align='center'}
# Write code here



```
