---
title: "Tidy Data"
subtitle: "EMSE 4575: Exploratory Data Analysis"
date: September 07, 2022
week: 2
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(readxl)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
```



# Practice 1: Reshaping Data

Run the code chunk to read in the following two data files:

- `pv_cell_production.xlsx`: Data on solar photovoltaic cell production by country
- `milk_production.csv`: Data on milk production by state

```{r}
milk_production <- read_csv(here::here('data', 'milk_production.csv'))
View(milk_production)

pv_cells <- read_excel(
    here::here('data', 'pv_cell_production.xlsx'),
    sheet = 'Cell Prod by Country',
    skip = 2
) %>%
    filter(!is.na(Year)) %>%
    # Need to convert a few columns so the data types match
    mutate(
        China = as.numeric(China),
        Taiwan = as.numeric(Taiwan),
        Malaysia = as.numeric(Malaysia),
        Germany = as.numeric(Germany),
        `South Korea` = as.numeric(`South Korea`),
        Others = as.numeric(Others)
    )
View(pv_cells)
```

Now modify the format of each:

- If the data are in "wide" format, convert it to "long" with `pivot_longer()`
- If the data are in "long" format, convert it to "wide" with `pivot_wider()`

```{r}
# Write code here

milk_wider <- milk_production %>% 
    pivot_wider(
        names_from=year,
        values_from = milk_produced
    )

View(milk_wider)

```

```{r}
# Write code here

pv_longer <- pv_cells %>% 
    pivot_longer(
        names_to= 'country', 
        values_to = 'num_cells',
        cols = -Year
    )

View(pv_longer)

```



```{r}
# Write code here

pv_longer_2 <- pv_cells %>% 
    pivot_longer(
        names_to= 'Country', 
        values_to = 'num_cells',
        cols = China:Others
    )

View(pv_longer_2)

```


```{r}
# Write code here

#pv_longer_3 <- pv_cells %>% 
#    pivot_longer(
#        names_to= 'Country', 
#        values_to = 'num_cells',
#        cols = c(China, Japan, Germany, Taiwan, Malaysia, `South Korea`) # c('China', 'Japan', 'Germany')
#    )
#
#View(pv_longer_3)

```

# Practice 2: Tidy Data Wrangling

Run the code chunk to read in the following two data files:

- `gapminder.csv`: Life expectancy in different countries over time
- `gdp.csv`: GDP of different countries over time

```{r}
gapminder <- read_csv(here::here('data', 'gapminder.csv'))
View(gapminder)

gdp <- read_csv(here::here('data', 'gdp.csv'))
View(gdp)
```

Now convert the data into a tidy (long) structure, then create the following summary data frames:

- Mean life expectancy in each year.
- Mean GDP in each year.


```{r}

# - Mean life expectancy in each year.

exp_per_country <- gapminder %>% 
    pivot_longer(
        names_to='Country',
        values_to='Life_Exp',
        cols = -year
    )

View(exp_per_country)

exp_per_year <- per_country %>%
  group_by(year) %>%
  summarise(Avg_Exp = mean(Life_Exp)) # %>%
  #arrange(avg_exp, desc = TRUE)

View(exp_per_year)
```



```{r}

# Mean GDP in each year

gdp_clean <- gdp %>% 
    pivot_longer(
        names_to='Year',
        values_to='GDP',
        cols = -country
    ) %>% 
    filter(!is.na(GDP)) # %>% 

View(gdp_clean)


gdp_per_year <- gdp_clean %>%
  group_by(Year) %>%
  summarise(Avg_GDP = mean(GDP)) # %>%

View(gdp_per_year)



#gdp %>% 
#    pivot_longer(
#        names_to = "year", 
#        values_to = "gdp", 
#        cols = -country
#    ) %>% 
#    filter(!is.na(gdp)) %>% 
#    group_by(year) %>% 
#    summarise(mean_gdp = mean(gdp))
```



# Practice 3: Tidy Data Visualization

Run the code chunk to read in the following two data files:

- `lotr_words.csv`: Word counts by character in the LOTR films
- `fed_spending.csv`: Federal R&D spending by department

```{r}
lotr <- read_csv(here::here('data', 'lotr_words.csv'))
fed_spending <- read_csv(here::here('data', 'fed_spend_wide.csv'))
View(fed_spending)
```

Now convert the data into a tidy (long) structure, then create the following charts

```{r}
fed_clean <- fed_spending %>% 
   pivot_longer(
        names_to='Dept',
        values_to='Budget_USD',
        cols = -year
    )  #%>%  

View(fed_clean)

# GROUP

budget_by_year <- fed_clean %>% 
    group_by(year) %>%
    summarise(Budget_Total = sum(Budget_USD))

View(budget_by_year)

# DATAVIZ
# https://www.geeksforgeeks.org/line-plot-using-ggplot2-in-r/

budget_by_year %>%
    ggplot(
        mapping = aes(x = year, y = Budget_Total)
    ) +
    geom_line() +
    geom_point() +
    labs(
      title = "Budget by Year",
      subtitle = "US Government Spending over time",
      caption = "This is a caption. Data source: ______________________.",
    ) +
    xlab("Year") +
    ylab("Budget Spending (USD)") +
    theme_classic()

```

