---
title: "Comparisons"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# install.packages("ggrepel")
# install.packages("ggridges")
library(ggrepel)
library(ggridges)
library(tidyverse)
library(here)
library(cowplot)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)

# Read in the data
gapminder        <- read_csv(here('data', 'gapminder.csv'))
milk_production  <- read_csv(here('data', 'milk_production.csv'))
internet_regions <- read_csv(here('data', 'internet_users_region.csv'))
```


### Your turn - reference lines

Use the `milk_production.csv` data to create the following charts showing differences from the mean state milk production in 2017.

```{r}
milk_summary <- milk_production %>%
  filter(year == 2017) %>%
  mutate(
    milk_produced = milk_produced / 10^9,
    state = fct_reorder(state, milk_produced))

View(milk_summary)
# looks like we have a row per state

avg_production = mean(milk_summary$milk_produced)
print(avg_production)
```

```{r milk-lollipop-mean, fig.height=8, fig.width=5.5, fig.align='center'}
# Write code here
```

```{r milk-bars, fig.height=8, fig.width=6, fig.align='center'}
# Write code here


# https://r-charts.com/part-whole/diverging-bar-chart-ggplot2/

milk_summary %>% 
    ggplot(
        aes(x = reorder(state, milk_produced), y = milk_produced)
    ) +
    geom_bar(stat = "identity", show.legend = FALSE) +
    #geom_vline(xintercept = avg_production, color = 'red', linetype = 'dashed') +
    geom_hline(yintercept = avg_production, color = 'red', linetype = 'dashed') +
    annotate(geom="text", label="The Mean", color="red", x="Georgia", y=5, hjust=0) +
    xlab("State") +
    ylab("Milk Produced") +
    coord_flip()


```



```{r milk-bars-diverging, fig.height=8, fig.width=6, fig.align='center'}

new_summary <- milk_summary
new_summary$mean_centered <- milk_summary$milk_produced - avg_production

new_summary <- new_summary %>% arrange(desc(milk_produced))

view(new_summary)
    

new_summary %>% 
    ggplot(
        aes(x = reorder(state, mean_centered), y = mean_centered)
    ) +
    geom_bar(stat = "identity", show.legend = FALSE) +
    #geom_vline(xintercept = avg_production, color = 'red', linetype = 'dashed') +
    geom_hline(yintercept = 0, color = 'red', linetype = 'dashed') +
    xlab("State") +
    ylab("Milk Produced") +
    coord_flip()


```






## Your turn - comparing multiple categories

Using the `internet_regions` data frame, pick a strategy and create an improved version of this chart.

Strategies:

- Dodged bars
- Facets
- Bullet chart
- Dumbell chart
- Slope chart

```{r internet-usage-bars, fig.height=6, fig.width=10, out.width="80%"}
View(internet_regions)

internet_regions_compare <- internet_regions %>%
  filter(year %in% c(2000, 2015)) %>%
  mutate(
    numUsers = numUsers / 10^6,
    year = as.factor(year))

View(internet_regions_compare)

ggplot(internet_regions_compare) +
  geom_col(
    aes(x = year, y = numUsers, fill = region),
    position = "dodge") +
  labs(y = "Millions of internet users")
```

```{r internet-usage-dodged, fig.height=6, fig.width=10, out.width="80%"}
# Make a dodged bar chart here
```

```{r internet-usage-facets, fig.height=4, fig.width=14, out.width="80%"}

#ggplot(internet_regions_compare) +
#    geom_col(
#        aes(x = year, y = numUsers, fill = region)
#    ) +
#    facet_wrap(vars(region), nrow = 1)
  
ggplot(internet_regions_compare) +
    geom_col(
        aes(y = year, x = numUsers, fill = region)
    ) +
    facet_grid(rows=vars(region))


ggplot(internet_regions_compare) +
    geom_col(
        aes(x = year, y = numUsers, fill = region)
    ) +
    facet_grid(cols=vars(region))


```

```{r internet-usage-bullet, fig.height=6, fig.width=10, out.width="80%"}
# Make a bullet chart here
```

```{r internet-usage-dumbbell, fig.height=6, fig.width=6, fig.align='center'}
# Make a dumbell chart here

# https://datavizpyr.com/dumbbell-plot-in-r-with-ggplot2/

internet_regions_compare %>% 
    ggplot(aes(x= numUsers, y= region)) +
    geom_line(aes(group = region)) +
    geom_point(aes(color=year), size=4) +
    theme(legend.position="right")


```

```{r internet-usage-slope, fig.height=7, fig.width=10, fig.align='center'}
# Make a slope chart here
```




## Your turn - comparing distributions

Use the `gapminder.csv` data to create the following charts comparing the distribution of life expectancy across countries in continents in 2007.

```{r}
gapminder_2007 <- gapminder %>%
  filter(year == 2007) %>%
  mutate(continent = fct_reorder(continent, lifeExp))
```

```{r gapminder-densities, fig.height=4.5, fig.width=6.5, fig.align='center'}
# Make an overlapping density chart here
```

```{r gapminder-ridges, fig.height=4.5, fig.width=6.5, fig.align='center'}
# Make an ridgeline chart here
```
