---
title: "Correlation"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    comment = "#>",
    fig.path = "figs/", # Plot save path
    fig.width = 7.252, # Plot dimensions
    fig.height = 4, #<<
    fig.retina = 3 # Better plot resolution
)

# install.packages('HistData')
# install.packages('palmerpenguins')
# install.packages('GGally')

library(HistData)
library(GGally)
library(palmerpenguins)
library(tidyverse)
library(here)

library(plotly) # :-)
 
# Set theme for all ggplots
theme_set(theme_bw(base_size = 20))

# Read in data
msleep <- read_csv(here('data', 'msleep.csv'))
```


```{r}
View(msleep)
View(penguins)
```

# 1. Correlation

Using the `penguins` data frame:

1. Find the two variables with the largest correlation in absolute value (i.e. closest to -1 or 1).

2. Create a scatter plot of those two variables.

3. Add an annotation for the Pearson correlation coefficient.

```{r}
# https://allisonhorst.github.io/palmerpenguins/

penguins %>% 
    ggcorr()

# LOOKS LIKE flipper_length_mm AND body_mass_g ARE MOST POSITIVELY CORRELATED

my_cor <- cor(penguins$flipper_length_mm, penguins$body_mass_g, method="spearman", use="complete.obs")
label_cor = paste("r=", round(my_cor,4))
print(label_cor)
#> 0.8399741

cor(penguins$flipper_length_mm, penguins$bill_depth_mm, method="spearman", use="complete.obs")
#> -0.5232675

```


```{r}
penguins %>% 
    select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g) %>% 
    ggpairs()
```


```{r}
# IN PYTHON THERE IS A PARAMETER TO DRAW THE TREND LINE BUT MAYBE NOT IN R :-/
plot_ly(y = penguins$flipper_length_mm, x = penguins$body_mass_g, type = "scatter") %>% 

```

```{r}
# USE GGPLOT :-D
# https://ggplot2.tidyverse.org/reference/annotate.html

penguins %>% 
    ggplot( 
        aes(x=flipper_length_mm, y = body_mass_g)
    ) + 
    geom_point() +
    annotate(geom = 'text',
             x = 215, # corresponds to the actual x and y values (not pixels)
             y = 3000, # corresponds to the actual x and y values (not pixels)
             label = label_cor,
             hjust = 0, 
             size = 7
    )

```

```{r}
penguins %>% 
    ggplot( aes(x=flipper_length_mm, y = body_mass_g)) + 
    geom_point() +
    geom_smooth(aes(x=flipper_length_mm, y = body_mass_g), method="lm", se=FALSE)
    annotate(geom = 'text',
             x = 215, # corresponds to the actual x and y values (not pixels)
             y = 3000, # corresponds to the actual x and y values (not pixels)
             label = label_cor,
             hjust = 0, 
             size = 7
    )

```


# 2. Linear models


```{r}
#model <- lm(formula = y ~ x, data = data)

model <- lm(flipper_length_mm ~ body_mass_g, data=penguins)
print(model)
model_coefs = coef(model) #> intercept, slope
print(model_coefs)
```



Using the `msleep` data frame:

1. Create a scatter plot of `brainwt` versus `bodywt`.
2. Include an annotation for the Pearson correlation coefficient.
3. Include an annotation for the best fit line.

Bonus: Compare your results to a log-linear relationship by converting the x and y variables to the log of x and y, like this:

```{r, eval=FALSE}
View(msleep)

x <- msleep$brainwt
y <- msleep$bodywt

model <- lm(log(x) ~ log(y), data = msleep)

model_coefs = coef(model)
print(model_coefs)

msleep %>% 
    ggplot( aes(x=brainwt, y = bodywt)) + 
    geom_point() +
    geom_smooth(aes(x=brainwt, y = bodywt), method="lm", se=FALSE)

#msleep %>% 
#    ggplot( aes(x=log(brainwt), y = log(bodywt))) + 
#    geom_point() +
#    geom_smooth(aes((x=log(brainwt), y = log(bodywt)), method="lm", se=FALSE)
#> You can also convert your plot to log axes by adding these layers:
#> plot +
#>     scale_x_log10() +
#>     scale_y_log10()

msleep %>% 
    ggplot( aes(x=brainwt, y = bodywt)) + 
    geom_point() +
    geom_smooth(aes(x=brainwt, y = bodywt), method="lm", se=FALSE) +
    scale_x_log10() +
    scale_y_log10()

```

```{r}
# https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/predict.lm

#
# GET EXAMPLE VALUES TO USE FOR PREDICTIONS
#

# test_set <- msleep %>% sample_n(1)
# print(test_set)
# x_test <- test_set$brainwt
# print(x_test) #> 0.169
# y_test <- test_set$bodywt
# print(y_test) #> 207.501
x_test <- 0.115
y_test <- 207.501

#y_pred = predict(model, x_test) 
##> data' must be a data.frame, environment, or list
y_pred = predict(model, data=c(x_test)) 
print(y_pred) #> a matrix WAT?

# SEE ALSO: https://www.tidymodels.org/
```





You can also convert your plot to log axes by adding these layers:

```{r, eval=FALSE}
plot +
    scale_x_log10() +
    scale_y_log10()
```

Linear model:

```{r}
# Write code here
```

Log-linear model:

```{r}
# Write code here
```

