---
title: "Trends"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# # Install new packages (only do this once!)
#install.packages('viridis')
#install.packages('gganimate')
# install.packages('gifski')
# install.packages('png')
#install.packages('magick')
# install.packages('av')

# FYI requires lower level dependency:
#  brew install imagemagick

library(knitr)
library(cowplot)
library(tidyverse)
library(viridis)
library(ggrepel)
library(gganimate)
library(here)
library(magick)

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
dplyr.width = Inf

# Load all the data sets:
global_temps <- read_csv(here::here('data', 'nasa_global_temps.csv'))
us_covid     <- read_csv(here::here('data', 'us_covid.csv'))

View(global_temps)
```

# 1. Visualizing single variables over time

Use the `global_temps` data frame to explore ways to visualize the change in average global temperatures.

Consider using:

- points
- lines
- areas
- smoothed lines

## Points

```{r global-temps-points, fig.height=5, fig.width=6}
# Write code here


global_temps %>% 
    ggplot() +
    geom_point(aes(x=year, y=meanTemp))


# LINEAR TREND LINE
global_temps %>% 
    ggplot(aes(x=year, y=meanTemp)) +
    geom_point() + 
    geom_smooth(method = "lm") +
    labs(title="Trendline: LM")


# BETTER TREND LINES
# https://ggplot2.tidyverse.org/reference/geom_smooth.html

global_temps %>% 
    ggplot(aes(x=year, y=meanTemp)) +
    geom_point() + 
    geom_smooth(method = "loess") +
    labs(title="Trendline: LOESS")


global_temps %>% 
    ggplot(aes(x=year, y=meanTemp)) +
    geom_point() + 
    geom_smooth(method = "gam") +
    labs(title="Trendline: GAM")
```

## Line

```{r global-temps-line, fig.height=5, fig.width=6}
# Write code here

global_temps %>% 
    ggplot() +
    geom_line(aes(x=year, y=meanTemp))

```

## Smoothed line

```{r global-temps-line-smoothed, fig.height=5, fig.width=6}
# Write code here
```

## Area

```{r global-temps-area, fig.height=5, fig.width=6}
# Write code here

global_temps %>% 
    ggplot() +
    geom_area(aes(x=year, y=meanTemp))


```

# 2. Animation

Use the `global_temps` data frame to animate one of the charts we did in the first exercise on changes in average global temperatures.

## Points

```{r global-temps-points-anim, fig.height=5, fig.width=6}


# https://gganimate.com/reference/renderers.html
# The gifski_renderer() is used unless otherwise specified in animate() or in options('gganimate.renderer').
# This renderer requires both the gifski and png packages to be installed.
#
# Other possible renderers are:
#
# magick_renderer() which requires the magick package and produce a gif. If gifski is not installed, the rendering will be much slower than using the gifski_renderer() and can potentially result in system problems when many frames need to be rendered (if gifski is installed magick will use it under the hood)


global_temps_anim <- global_temps %>% 
    ggplot() +
    geom_point(aes(x=year, y=meanTemp)) + 
    transition_reveal(year) + 
    shadow_trail(0.01) + # Leave a trail of the points
    labs(title = "Year: {frame_time}") # not working dynamically?


animate(global_temps_anim,
        end_pause = 15,
        duration = 10,
        width = 1100, height = 650, res = 150,
        renderer = magick_renderer()
)

# animation: The animation object to save, as returned by animate(). Defaults to the last rendered animation using last_animation()
filepath <- here::here('global_temps_animation.gif')
anim_save(filepath)

```

## Line

```{r global-temps-line-anim, fig.height=5, fig.width=6}
# Write code here
```

## Area

```{r global-temps-area-anim, fig.height=5, fig.width=6}
# Write code here
```

# 3. Visualizing mutliple variables over time

Use the `us_covid` data frame to explore ways to visualize the number of daily cases using:

1. A labeled line chart
2. A stacked area chart
3. A heat map


```{r}
View(us_covid)

us_covid %>% ggplot() +
    geom_line(aes(x=date, y=cases_daily))


#us_covid %>% 
#    ggplot(aes(x=date, y=cases_daily)) +
#    geom_point() +
#    geom_smooth(method = "loess")
#
```

## Labeled line chart

For this one, try showing just the lines for NY and CA.
I already made the summary data frame to do this.

```{r covid_ca_ny, echo=FALSE, fig.height=4, fig.width=12, fig.align='center'}
us_covid_ca_ny <- us_covid %>%
  mutate(cases_daily = cases_daily / 10^3) %>%
  filter(state %in% c("California", "New York"))

# Write ggplot code here
```

## Stacked area chart

For this one, try showing NY, CA, and all other states.
I already made the summary data frame to do this.

```{r covid_area, fig.height=4, fig.width=12}
us_covid %>%
  mutate(
    state = fct_other(state, keep = c(
      "California", "New York")),
    state = fct_relevel(state, c(
      "Other", "California", "New York"
    ))) %>%
  group_by(day, state) %>%
  summarise(deaths_daily = sum(deaths_daily))

# Write ggplot code here
```

## Heat map

```{r covid_heatmap, fig.height=7, fig.width=10}

# https://r-graph-gallery.com/79-levelplot-with-ggplot2.html

#us_covid %>% ggplot() +
#    geom_tile(aes(x=date, y=state, fill=cases_daily))

#us_covid %>% ggplot() +
#    geom_tile(aes(x=date, y=state, fill=cases_daily)) +
#    scale_fill_distiller(palette = "RdPu") #+
#    #theme_ipsum() #> Error in theme_ipsum() : could not find function "theme_ipsum"

us_covid %>% ggplot() +
    geom_tile(aes(x=date, y=state, fill=cases_daily)) +
    scale_fill_gradient(low="white", high="red")





```
