---
title: "Centrality & Variability"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    comment = "#>",
    fig.path = "figs/", # Plot save path
    fig.width = 7.252, # Plot dimensions
    fig.height = 4, #<<
    fig.retina = 3 # Better plot resolution
)

theme_set(theme_bw(base_size = 20)) # Set theme for all ggplots

# Read in data
milk_production <- read_csv(here('data', 'milk_production.csv'))
lotr_words <- read_csv(here('data', 'lotr_words.csv'))
marathon <- read_csv(here('data', 'marathon.csv'))
wildlife_impacts <- read_csv(here('data', 'wildlife_impacts.csv'))
faithful <- read_csv(here('data', 'faithful.csv'))
candy_rankings <- read_csv(here('data', 'candy_rankings.csv'))
```

# Quick practice: What's the data type?

Decide here:

https://docs.google.com/presentation/d/1J8UtyEwkA5QEcQQ9LCAs4EU1gyhPY7QIoIW_T_gys6o/edit?usp=sharing

```{r}
#wildlife_impacts %>%
#  filter(!is.na(cost_repairs_infl_adj)) %>%
#  select(incident_date, time_of_day, species, cost_repairs_infl_adj)

wildlife_clean <- wildlife_impacts %>%
  filter(!is.na(cost_repairs_infl_adj)) %>%
  select(incident_date, time_of_day, species, cost_repairs_infl_adj)

View(wildlife_clean)
glimpse(wildlife_clean)

typeof(wildlife_clean) #> list
typeof(wildlife_clean$incident_date)
typeof(wildlife_clean$time_of_day)
typeof(wildlife_clean$species)
typeof(wildlife_clean$cost_repairs_infl_adj)
```

- incident_date: double
- time_of_day: character
- species: character
- cost_repairs_infl_adj: double



# Practice with summary measurements

1) Read in the following data sets:

- milk_production.csv
- lotr_words.csv

 2) For each variable in each data set, if possible, summarize its

- Centrality
- Variability

```{r}
# milk_production:

View(milk_production)

#summarise(milk_production) # NOPE
summary(milk_production) # YEP
#>   region             state                year      milk_produced
#> Length:2400        Length:2400        Min.   :1970   Min.   :3.000e+06
#> Class :character   Class :character   1st Qu.:1982   1st Qu.:4.610e+08
#> Mode  :character   Mode  :character   Median :1994   Median :1.348e+09
#>                                       Mean   :1994   Mean   :3.132e+09
#>                                       3rd Qu.:2005   3rd Qu.:2.744e+09
#>                                       Max.   :2017   Max.   :4.234e+10

milk_production %>%
    summarise(
      mean = mean(milk_produced), #> 3_131_641_667
      median = median(milk_produced), #> 1_347_500_000
      stdev = sd(milk_produced), #> 5_422_560_750
      iqr = IQR(milk_produced) #> 2_283_250_000
    )

# BY REGION

milk_production %>%
    group_by(region) %>%
    summarise(
      mean = mean(milk_produced), #> 3_131_641_667
      median = median(milk_produced), #> 1_347_500_000
      stdev = sd(milk_produced) #> 5_422_560_750
    ) %>%
    arrange(desc(mean))

# BY STATE

milk_production %>%
    group_by(state) %>%
    summarise(
      mean = mean(milk_produced), #> 3_131_641_667
      median = median(milk_produced), #> 1_347_500_000
      stdev = sd(milk_produced) #> 5_422_560_750
    ) %>%
    arrange(desc(mean)) # https://dplyr.tidyverse.org/reference/desc.html
```

```{r}

# lotr_words

View(lotr_words)

lotr_words %>%
    summarise(
      mean_male = mean(Male),  #> 2_090.778
      median_male = median(Male), #> 2_459
      stdev_male = sd(Male), #> 1_200.084
      mean_female = mean(Female),  #> 269.7778
      median_female = median(Female), #> 183
      stdev_female = sd(Female), #> 392.3161
    )

```



# Practice with visual summaries

1) Read in the following data sets:

- faithful.csv
- marathon.csv

2) Summarize the following variables using an appropriate chart (bar chart, histogram, and / or boxplot):

- faithful: eruptions
- faithful: waiting
- marathon: Age
- marathon: State
- marathon: Country
- marathon: `Official Time`

```{r}
# https://plotly.com/r/histograms/
# https://plotly.com/r/reference/#histogram
# https://stackoverflow.com/questions/54816608/customizing-bin-widths-in-plotlys-histogram-function-in-r

library(plotly)

# faithful: eruptions
View(faithful)

fig <- plot_ly(x = faithful$eruptions, type = "histogram", nbinsx=30)
fig

```



```{r}

# faithful: waiting

plot_ly(x = faithful$waiting, type = "histogram", nbinsx=30)

```


```{r}
# marathon: Age

View(marathon)

plot_ly(x = marathon$Age, type = "histogram", nbinsx=100)
```



```{r}

# marathon: State

# let's count the number of participants from each state
# https://dplyr.tidyverse.org/reference/count.html

runners_by_state <- marathon %>%
    group_by(State) %>%
    summarise(
        #runner_count=count() # NOPE
        #runner_count= tally() # NOPE
        runner_count= n()  # YES!
    ) %>%
    arrange(desc(runner_count))

View(runners_by_state)

#plot_ly(x = runners_by_state$State, y= runners_by_state$runner_count, type = "bar")
#
#plot_ly(y = runners_by_state$State, x= runners_by_state$runner_count, type = "bar", orientation="h")
#
#plot_ly(y = runners_by_state$State, x= runners_by_state$runner_count, type = "bar", orientation="h") %>%
#layout(xaxis = list(categoryorder = "total descending"))
#%>%
#    layout(
#         title = "Runners by state",
#         xaxis = list(title = "X AXIS LABEL", categoryorder = "array", categoryarray = runners_by_state$runner_count),
#         yaxis = list(title = "Y AXIS LABEL")
#)

plot_ly(x = runners_by_state$State, y= runners_by_state$runner_count, type = "bar") %>%
    layout(
        title="Runners by state",
        xaxis= list(categoryorder= "total descending")
    )

plot_ly(y = runners_by_state$State, x= runners_by_state$runner_count, type = "bar", orientation="h") %>%
    layout(
        title="Runners by state",
        yaxis= list(title="State", categoryorder= "total ascending"),
        xaxis= list(title="Number of Runners")
    )

```



```{r}

# marathon: Country



```


```{r}


# marathon: `Official Time`

```








# Practice with visualizing _relationships_

1) Read in the following data sets:

- marathon.csv
- wildlife_impacts.csv

2) Visualize the _relationships_ between the following variables using an appropriate chart (bar plots, scatterplots, and / or box plots):

- marathon: Age & `Official Time`
- marathon: Country & `Official Time`
- wildlife_impacts: state & operator

```{r}
# marathon: Age & `Official Time`

View(marathon)
summary(marathon$`Official Time`)

times_by_age <- marathon %>% 
    group_by(Age) %>% 
    summarize(avg_time = mean(`Official Time`))
    
View(times_by_age)
summary(times_by_age)
# hmm the times are formatted as 'X sec' (string) hmmmmm

plot_ly(x = times_by_age$Age, y= times_by_age$avg_time, type = "bar") %>%
    layout(
        title="Average Finish Time by Age",
        xaxis= list(title="Age"),
        yaxis= list(title="Mean Finish Time")
    )
```

```{r}

# marathon: Country & `Official Time`



```

```{r}

# wildlife_impacts: state & operator


```


# Practice doing EDA

1) Read in the candy_rankings.csv data sets

2) Preview the data, note the data types and what each variable is.

3) Visualize (at least) three _relationships_ between two variables (guided by a question) using an appropriate chart:

- Bar chart
- Scatterplot
- Boxplot

```{r}

```
